GVN-Hoist: Hoisting Computations from Branches

### 2016-08-17 hoist.tex
grammar rewriting

\subsection{Finding the candidates}
Finding if a set of instructions perform identical computations. Since this code
hoisting is based on GVN, we need the facility to compute GVN of instructions.
In llvm \cite{llvm}, GVN infrastructure is already set up and works well for
most kind of instructions except loads and stores so we compute the GVN of loads
and stores separately. In order to value number loads, we value number (hash)
the address from where the value is to be loaded. For stores, we value number
(hash) the address as well as the value to be stored at that address.

@@

\subsection{Checking for legality}
\label{subsec:legality}
Finding a hoisting point i.e., a common dominator of the set of instructions. If
such a dominator is found then checking if all the use-operands of the set of
instructions are available or not. In some cases when the operands are not
available, it is possible to reinstantiate (remateralize) the use-operands, thus
passing the legality check. Checking that the side-effects of the computations
does not intersect with any side-effects between the instructions to be hoisted
and their hoisting point. For memory operations like loads, stores, calls etc.,
it is also required to check that all the paths from the hoisting point to the
end of the function should execute the exact instruction, in order to guarantee
correctness. Moreover, hoisting of memory operations is tricky on paths which
have indirect branch targets e.g., landing pad, case statements, goto labels
etc., because it becomes difficult to prove that all the paths from hoisting
point to the end of the function would execute the instruction. In our current
implementation we discard hoisting through such paths.

@@

\subsubsection{Hoisting scalars}
Scalars are the easiest to hoist because we do not have to analyze them for
aliasing memory references. As long as all the operands are available, the
scalar can be hoisted.

>>>>

\subsection{Finding the candidates}
The first step is to find if a set of instructions perform identical
computations. Since our code hoisting algorithm is based on GVN, we need the
facility to compute GVN of instructions.  In llvm \cite{llvm}, GVN
infrastructure is already set up and works well for most kind of instructions
except loads and stores so we compute the GVN of loads and stores separately. In
order to value number loads, we value number (hash) the address from where the
value is to be loaded. For stores, we value number (hash) the address as well as
the value to be stored at that address.

@@

\subsection{Checking for legality}
\label{subsec:legality}
Since the equality of candidates is purely based on the value they compute, we
need to establish if hoisting them to a common dominator would be feasible. One
a common dominator is found then we check if all the use-operands of the set of
instructions are available or not. In some cases when the operands are not
available, it is possible to reinstantiate (remateralize) the use-operands, thus
passing the legality check. Subsequently, it is checked that the side-effects of
the computations does not intersect with any side-effects between the
instructions to be hoisted and their hoisting point. For memory operations like
loads, stores, calls etc., it is also required to check that all the paths from
the hoisting point to the end of the function should execute the exact
instruction, in order to guarantee correctness. Moreover, hoisting of memory
operations is tricky on paths which have indirect branch targets e.g., landing
pad, case statements, goto labels etc., because it becomes difficult to prove
that all the paths from hoisting point to the end of the function would execute
the instruction. In our current implementation we discard hoisting through such
paths.

@@

\subsubsection{Hoisting scalars}
Scalars are the easiest to hoist because we do not have to analyze them for
aliasing memory references. As long as all the operands are available (or can be
made available by rematerialization), the scalar can be hoisted.
